Spatial Reasoning and Statistical Graphics

### 2014-11-03 Paper/Draft1/TestingVisualAptitude.Rnw
couple of comments for PCA section

<<ResultsScaledScores, echo=F, include=T, fig.pos="hbt", fig.cap="Test scores for lineups and visuospatial tests. As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing. We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability. \\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=

@@

Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are measuring visual ability of individuals in some way. 
%are influenced by factors other than visual ability -- general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few. 
What is of more interest to us, is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores. 
\comment{give a mini-paragraph summary of the PCA of four}
%
Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance. 
PC1
accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor. %, that is, all of the general factors discussed above which influence test results. 
We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
Biplots of these principal components are shown in figure~\ref{fig:biplots5}.

>>>>

<<ResultsScaledScores, echo=F, include=T, fig.pos="htb", fig.cap="Test scores for lineups and visuospatial tests. As none of the participants scored at or below zero, we can conclude that there is little evidence of random guessing. We also note the score compression that occurs on the Visual Search test; this indicates that most participants scored extremely high, and thus, participants' scores are not entirely representative of their ability. \\label{fig:Scores}", fig.width=4.5, fig.height=4.5, out.width='\\linewidth', warning=F, message=F>>=

@@

Results from the visuospatial tests used in this experiment are highly correlated, as shown in figure~\ref{fig:scatterplotmatrix}; this is to be expected given that all of these tests are measuring visual ability of individuals in some way. 
%are influenced by factors other than visual ability -- general intelligence, mental processing speed, cognitive resources, motivation, and attention, to name a few. 
What is of more interest to us, is how other factors, such as e.g.~general intelligence, mental processing speed, cognitive resources, motivation, and attention affect performance. 
In order to assess factors contributing to lineup performance, we first examine the separate dimensions measured by the entire battery of tests using principal components analysis on the scaled test scores. 
\comment{give a mini-paragraph summary of the PCA of four}
%
\newtext{A principal component analysis (PCA) of the four established visuo-spatial tests reveals that they all share a very strong first component, which explains about XXX\% of the total variability. PC1 is essentially an average across all tests representing a general ``visual intelligence" factor. The principal components beyond that span another two dimensions, while the last dimension is weak (at XXX\%). The second principal component differentiates the figure classification task from card rotation, PC3  ... PC4 ...}

\comment{then move on to the PCA of five, but summarize strongly - what does this one have in common with the PCA of four. The comments from below would be great here.}
Table~\ref{tab:PCAvariance5} shows the importance of each principal component and the cumulative proportion of variance. 
PC1
accounts for \Sexpr{round(summary(pca)$importance[3,1]*100,1)}\% of the variance in the data, the corresponding rotation matrix is provided in~\ref{tab:PCArotation5}.
As expected, PC1 is essentially an average across all tests representing a general ``visual intelligence" factor. %, that is, all of the general factors discussed above which influence test results. 
We will focus on the remaining components, which show differences in performance on the tests, and thus provide us with more insight about the skills needed to successfully examine lineups. 
Biplots of these principal components are shown in figure~\ref{fig:biplots5}.
